```{r echo=FALSE, message=FALSE}
setwd("..")
require(ProjectTemplate)
load.project()
theme_set(theme_bw())
```

Home Ice Advantage in College Hockey
====================================

As a Cornell hockey fan, I've often heard Cornell's home rink, Lynah Rink, described as a tough place for opposing teams to play. Recently, I've started to wonder if that's actually true, and if so, what it means. Is Lynah a tough place to play because you have to play Cornell there, and they have typically had strong teams? Or does Cornell have  a particularly strong home ice advantage? More generally, does home ice advantage exist in college hockey, and if so, how is it characterized?

This is not intended as a very technical analysis. In fact, is as much an exercise in learning to scrape data from the web, use RStudio, and produce a report using R Markdown. But I hope to address a handful of questions about home ice advantage in college hockey. Specifically:
* How should home ice advantage be measured?
** Better winning percentage at home?
** More goals scored at home?
** Fewer goals allowed at home?
** Higher goal differential at home?
* Does home ice advantage exist?
* Is the overall level of home ice advantage across college hockey stable over time?
* Are individual teams' home ice advantages stable over time?

Measuring Home Ice Advantage
----------------------------

Does Home Ice Advantage Exist?
------------------------------

The answer to this question is a resounding YES! Over the past fifteen seasons (1998-2013), `r sprintf("%0.1f", 100*mean(agg.data$goal.diff.adv > 0))`% of team-seasons experienced a higher goal differential at home than on the road. Over the entire fifteen year period, every single team has had a positive aggregate goal differential advantage. The following histogram shows the distribution of team aggreagate advantage over the sample period.

```{r echo=FALSE, fig.width=7, fig.height=6}
temp <- agg.data[, list(adv = sum(games*goal.diff.adv)/sum(games)), by = list(team)]
print(ggplot(temp, aes(adv)) 
      + geom_histogram(binwidth=0.25, fill="lightblue", color="black")
      + xlab("Goal Differential Advantage") + ylab("Count")
      + ggtitle("Histogram of Team Aggregate Goal Differential Advantage"))
```

The following plot shows the distribution of goal differential advantage for each team over the sample period.

```{r echo=FALSE, fig.width=7, fig.height=6}
temp <- agg.data[, list(m = median(goal.diff.adv)), by = list(team)]
teams.sorted <- temp$team[order(-temp$m)]
print(ggplot(agg.data, aes(factor(team, levels=teams.sorted), goal.diff.adv)) 
      + geom_boxplot(fill="lightblue") 
      + theme(axis.text.x = element_text(angle=90, hjust=1))
      + xlab("Team") + ylab("Goal Differential Advantage")
      + ggtitle("Boxplots of Team Goal Differential Advantage Across Seasons"))

```

The following plot shows the correlation of team winning percentage and goal differential advantage by season.

```{r echo=FALSE, fig.width=7, fig.height=6}
temp <- agg.data[, list(corr = cor(win.pct, goal.diff.adv)), by = list(season)]
print(ggplot(temp, aes(season, corr)) 
      + geom_line() + geom_point()
      + ggtitle("Correlation of Win Pct. and Goal Diff. Adv.") 
      + xlab("Season") + ylab("Corr(Win Pct., Goal Diff. Adv."))
```